---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

# bysykkel

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/bysykkel)](https://cran.r-project.org/package=bysykkel)
[![license](https://img.shields.io/github/license/mashape/apistatus.svg?maxAge=2592000)](https://github.com/PersianCatsLikeToMeow/bysykkel/blob/master/LICENSE)
<!-- badges: end -->

## Overview

bysykkel provides functions to easily process Norwegian city bike data, either 
by reading it directly to R or downloading it to your R session's working 
directory.

* `read_trips_data()` reads bike trip data to R as a dataframe.
* `fread_trips_data()` fast reads bike trip data to R as a dataframe by 
utilizing `fread()` from data.table.
* `dl_trips_data()` downloads bike trip data to your working directory as a 
CSV-file.
* `get_api_data()` gets realtime data from the city bike API service.

bysykkel lets you, the user, focus on data exploration, visualization, 
statistical analysis, and building machine learning models on Norwegian
city bike data, by simplifying the task of getting the data. Indeed, the
purpose of bysykkel is to minimize time spent on getting Norwegian city 
bike data, and lower the barrier to start analyzing the data.

The package name is the Norwegian word for "city bikes", where *by* means 
"city", and *sykkel* means "bike" (or "bicycle").

## Installation

You can install the released version of bysykkel from 
[CRAN](https://CRAN.R-project.org) with:

```{r, eval = FALSE}
install.packages("bysykkel")
```

### Development version

Alternatively, you can install a development version of bysykkel from GitHub
to get bug fixes or new features before the next package version is released 
on CRAN. To install the development version, you can use devtools to install
bysykkel from GitHub.

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("PersianCatsLikeToMeow/bysykkel")
```

## Bike data

bysykkel interacts with four city bike services in Norway that make 
bike data publicly available[^1]:

[^1]: Bike data for
[Bærum City Bike](https://www.baerum.kommune.no/tjenester/vei-trafikk-og-parkering/sykkel-i-barum/bysykkel/) 
(unavailable in English) and 
[Oslo Cargo Bike](https://oslolastesykkel.no/en)
is not publicly available. This 
[article](https://medium.com/urbansharing/piloting-cargo-bikes-to-study-oslos-combined-mobility-needs-b4a8bf536c60)
states that "(...) all data and
findings wil be shared openly (...)" regarding the Oslo Cargo Bike pilot
in collaboration with Hertz's carpool service. However, that data is either
(1) available somewhere else than the Oslo Cargo Bike website; or (2) not yet
publicly available.

* [Oslo City Bike](https://oslobysykkel.no/en/open-data)
* [Oslo Winter Bike](https://oslovintersykkel.no/en/open-data)
* [Bergen City Bike](https://bergenbysykkel.no/en/open-data)
* [Trondheim City Bike](https://trondheimbysykkel.no/en/open-data)

Each city bike service provide two data-related services:

* Historical trip data
* Realtime data

The historical trip data is available both as a CSV-file, and a JSON-file,
that contains monthly anonymized historical bike trip data. Realtime
data is available in the 
[GBFS format](https://github.com/NABSA/gbfs/blob/master/gbfs.md), and must
be accessed with each city bike's API service, which provide information
about 

* bike and dock availability;
* stations (including geolocation, addresses and station descriptions); and
* machine readable basic information about the city bike service. 

The data is made available under the Norwegian License for Open Data 2.0,
abbreviated as NLOD 2.0, which you can read about 
[here](https://data.norge.no/nlod/en/2.0).

## Examples

### Read bike trips data to R

```{r, eval = FALSE, message = FALSE}
library(bysykkel)

# Get bike trip data for April, 2019 for Oslo as a dataframe
oslo.bike <- read_trips_data(year = 2019, month = 04, city = "Oslo")

# Get winter bike data for January, 2019 for Oslo as a dataframe
oslo.winter.bike <- read_trips_data(2019, 1, "OsloW")

# Fast read bike data from June to August in 2018 for Bergen with `lapply()`,
# and `rbind()` the resulting `list` with `do.call()` to get a dataframe

#! install.packages("data.table")

bergen.bike <- lapply(06:08, fread_trips_data, year = 2018, city = "Bergen")
bergen.bike <- do.call(rbind, bergen.bike)

# Alternatively, use `map_dfr()` from `purrr` instead of `lapply()`,
# `rbind()`, and `do.call() to get the same result: a dataframe

#! install.packages("purrr")

library(purrr)

bergen.bike <- map_dfr(6:8, fread_trips_data, year = 2018, city = "Bergen")
```

**NB!** I recommend that you use `fread_trips_data()` to fast read city bike 
data, especially if you want to read bike data for multiple months.

**NB!** data.table is not automatically installed with bysykkel, and must be
installed separately with `install.packages("data.table")` if you want to
use `fread_trips_data()`.

### Download bike trips data

```{r, eval = FALSE, message = FALSE}
library(bysykkel)

# Download bike trip data for April 2019 for Trondheim
dl_trips_data(2019, 04, "Trondheim")
#> The CSV-file is downloaded to your R session's current working directory

# Download bike trip data for summer 2018 for Oslo
lapply(06:08, dl_trips_data, year = 2018, city = "Oslo")
#> The CSV-file for each month is downloaded to your R session's working directory
```

### Get realtime data from the API service

**NB!** Please read each City Bike's guide on how to correctly use their API 
service before using `get_api_data()`. See
[Oslo City Bike's guide](https://oslobysykkel.no/en/open-data/realtime)
as an example.

The `return_df` argument in `get_api_data()` specifies whether you want to
return the result as a list. If `return_df = FALSE`, then the function 
returns a list that contains a dataframe, and a number that represents the
(POSIX) time of when you made the API request.

```{r, eval = FALSE, message = FALSE}
library(bysykkel)

# Get API data on bike stations as a dataframe
oslo.stations <- get_api_data(client_id = "myname-myapp", 
                              data = "stations",
                              return_df = TRUE)

# Get API data for bike availability as a list that contains a dataframe, and
# a number that represents the (POSIX) time of when you made the API request
bergen.availability <- get_api_data(client_id = "mycompany-myservice",
                                    data = "availability",
                                    return_df = FALSE)

# Get API data on bike system information
trondheim.system.information <- get_api_data("Ola Nordmann-bike dashboard",
                                             "system",
                                             return_df = FALSE)
```

## File an issue or suggest an improvement

If you want to report a discovered bug, raise some other issue, or suggest an
improvement to bysykkel, then please file an issue on [GitHub](https://github.com/PersianCatsLikeToMeow/bysykkel/issues).
For bugs, please file a minimal reproducible example.

---
